---
title: "introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 6
)
library(syntheticnetworks)
```

# Introduction

The **syntheticnetworks** package provides algorithms to simulate spatial
distributions of habitat patches and calculating transition matrices based on
dispersal kernels.

Currently, four algorithms are provided for generating spatial patch
configurations:

* **Uniform**

* **Clumpy**

* **Lacy**

* **Stringy**

This vignette introduces each algorithm with runnable examples.

## Uniform patches

Uniformly distributed patches are generated by sampling random $(x, y)$
coordinates across the landscape.

```{r setup}

uniform <- generate_uniform_patches(n.patches = 200, extent = 10)

# Visualise
plot(uniform$x, uniform$y, pch = 19, col = "blue",
     main = "Uniform patches", xlab = "x", ylab = "y")

```

## Clumpy patches

Clumpy patches are generated around random foci with a specified mean clump
length and clump size.

```{r setup}

# Generate a clumpy patch landscape
patches <- generate_clumpy_patches(n.patches = 200, extent = 10, clump_length = 1,
                                   n.foci = 5)

# Visualise
plot(patches$x, patches$y, main = "Clumpy Patches", pch = 19)

# Create transition matrix
tm <- calculate_transition_matrix(patches, dispersal = 2, R = 1)
```
